Quantum Walks and Applications to Quantum Money

### 2025-03-29 presentation.tex
new commit

\begin{frame}{Program and Courses Taken}

    \begin{textblock*}{12cm}(0.5cm,1.5cm)
       \begin{itemize}
        \item Program: M.A.Sc in Software Engineering
        \vspace{1cm}
        \item CAS 701,  Logic \& Discrete Mathematics
        \item COMPSCI 6TE3, Continuous optimization
        \item CAS 721, Combinatorics \& Computing
        \item CAS 741, Development of Scientific Computation Software
       \end{itemize}
    \end{textblock*}

    \title{SDSD}

\begin{frame}{Seminars}

    \begin{textblock*}{12cm}(0.5cm,1.5cm)
       \begin{itemize}
        \item I have participated in 6 seminars during my program.
       \end{itemize}
    \end{textblock*}

\begin{frame}{Poster}

    \begin{figure}
        \centering
        \includegraphics[width=1\textwidth]{poster.png}
        \caption{Presented at Nov. 6, 2024}
    \end{figure}

@@

\begin{frame}{The Hartley Transform}

    \begin{textblock*}{12cm}(0.5cm,1.5cm)
        item Let $N$ be a positive integer, and let $\mathbb{Z}_N$ be the additive cyclic group of integers modulo $N$. The Hartley transform of a function $f: \mathbb{Z}_N \to \mathbb{R}$ is the function $\cht_N(f): \mathbb{Z}_N \to \mathbb{R}$ defined by
        \[ \cht_N(f)(a) = \frac{1}{\sqrt{N}} \sum_{y = 0}^{N - 1} \cas\Big(\frac{2 \pi ay}{N}\Big) f(y),  \]
        where $\cas(x) = \cos(x) + \sin(x)$ \\
        For a single basis element of the cyclic group $\mathbb{Z}_N$, the quantum Hartly transform simplifies to
        \begin{equation}
            \label{eq:qht-N}
            \qht_N: \ket{a} \mapsto \frac{1}{\sqrt{N}} \sum_{y = 0}^{N - 1} \cas\Big(\frac{2\pi a y}{N}\Big) \ket{y}.
        \end{equation}
    \end{textblock*}

@@

    \begin{textblock*}{12cm}(0.5cm,1.5cm)
        First, let us briefly explain how the algorithm for $\qft_N$ works:
        \begin{align*}
            \qft_N\ket{a}
            & = \frac{1}{\sqrt{N}} \sum_{y = 0}^{N - 1} \omega_N^{ay}\ket{y} \\
            & = \frac{1}{\sqrt{N}} \sum_{y = 0}^{N / 2 - 1} \omega_N^{ay} \ket{y} + (-1)^a \sum_{y = 0}^{N / 2 - 1} \omega_N^{ay} \ket{y + N/2} \\
            & = \frac{1}{\sqrt{N / 2}} \sum_{y = 0}^{N / 2 - 1} \omega_N^{ay} \frac{1}{\sqrt{2}} (\ket{0} + (-1)^a \ket{1}) \ket{y}, \numberthis\label{eq:qft_alt}
        \end{align*}

@@

@@

        The adjacency matrix of the Cayley graph $\Gamma = (V,E)$ can be expressed as:
        \[ A = \sum_{a \in G} \lambda_a \ket{\hat{a}} \bra{\hat{a}}, \]
        where $\ket{\hat{a}} = \operatorname{QFT} \ket{a}$ is the quantum Fourier transform of $\ket{a}$. \\
         \vspace{0.5cm}
        The eigenvalues $\lambda_a$ are given by:
        \[ \lambda_a = \sum_{q \in Q} \chi(a, q). \]
        Note that the eigenvectors $\ket{\hat{a}}$ of $A$ depend only on $G$ and not on the set $Q$.

    Given a regular group action $(G, X, *)$ with a fixed element $x \in X$ and a set  $Q = \{q_1, q_2, \dots, q_k\} \subset G$, let $\Gamma = (X, E)$ be a graphs with vertex set $X$ and edge set consisting of pairs $(x, y) \in X \times X$ such that $y = q * x$ for some $q \in Q$. The adjacency matrix of $\Gamma$ is
    \[
    A = \sum_{h \in G} \lambda_h \ket{G^{(h)} * x} \bra{G^{(h)} * x}, 
    \]
    where $\lambda_h = \sum_{q \in Q} \chi(h, q)$. Again, the eigenvectors $\ket{G^{(h)} * x}$ depend only on $G$.

        \[
        e^{-iAt} = \sum_{h \in G}\ket{G^{(h)}*x}\bra{G^{(h)}*x}
        \]

        \[
        T := \sum_{y \in X}\sum_{b \in {\{0,1\}}}(\ket{j}\bra{j}\otimes\ket{b}\bra{b})\otimes\ket{\phi_{yb}}
        \]

        \[
        = \sum_{y \in X}\sum_{b \in {\{0,1\}}}\ket{y}\ket{b}\ket{\phi_{yb}}\bra{y}\bra{b}
        \]

        \[
            \ket{\phi_{y0}} = \frac{1}{\sqrt{|Q|}}\sum_{q \in Q}\ket{q*y}\ket{0}
        \]

        Proof. 
        \begin{align*}
            e^{iAt} \ket{\mathbb{Z}_N^{(h)} * x}_H
            & = \sum_{g \in \mathbb{Z}_N} e^{i\lambda_gt} \ket{\mathbb{Z}_N^{(g)} * x}\braket{\mathbb{Z}_N^{(g)} * x|\mathbb{Z}_N^{(h)} * x}_H \\
            & = \sum_{g \in \mathbb{Z}_N} e^{i\lambda_gt} \ket{\mathbb{Z}_N^{(g)} * x} \bra{\mathbb{Z}_N^{(g)} * x} \Big( \frac{1 - i}{2} \ket{\mathbb{Z}_N^{(h)} * x} + \frac{1 + i}{2} \ket{\mathbb{Z}_N^{(-h)} * x} \Big) \\
            & = e^{i\lambda_{h}t} \frac{1 - i}{2} \ket{\mathbb{Z}_N^{(h)} * x} + \frac{1 + i}{2} e^{i\lambda_{-h}t} \ket{\mathbb{Z}_N^{(-h)} * x} \\
            & = e^{i\lambda_{h}t} \ket{\mathbb{Z}_N^{(h)} * x}_H,
        \end{align*}
        where the last equality follows from the fact that $\lambda_h = \lambda_{-h}$.
    \end{textblock*}

        \vspace{1cm}
        If we choose $t = \operatorname{poly}(\log N)$, it follows from Lemma that we can run the phase estimation algorithm with the unitary $W$ and the eigenstate $\ket{\mathbb{Z}_N^{(h)} * x}_H$ to compute an estimate $\tilde{\lambda}_h$ of $\lambda_h$ such that %$\abs{\tilde{\lambda}_h - \lambda_h} \le 1 / poly(log N)$. 
        % \operatorname{poly}(\log N)
        \[
        |\tilde{\lambda}_h - \lambda_h| \leq \frac{1}{\operatorname{poly}(\log N)}
        \]

>>>>

% \begin{frame}{Program and Courses Taken}

%     \begin{textblock*}{12cm}(0.5cm,1.5cm)
%        \begin{itemize}
%         \item Program: M.A.Sc in Software Engineering
%         \vspace{1cm}
%         \item CAS 701,  Logic \& Discrete Mathematics
%         \item COMPSCI 6TE3, Continuous optimization
%         \item CAS 721, Combinatorics \& Computing
%         \item CAS 741, Development of Scientific Computation Software
%        \end{itemize}
%     \end{textblock*}

%     \title{SDSD}

% \begin{frame}{Seminars}

%     \begin{textblock*}{12cm}(0.5cm,1.5cm)
%        \begin{itemize}
%         \item I have participated in 6 seminars during my program.
%        \end{itemize}
%     \end{textblock*}

% \begin{frame}{Poster}

%     \begin{figure}
%         \centering
%         \includegraphics[width=1\textwidth]{poster.png}
%         \caption{Presented at Nov. 6, 2024}
%     \end{figure}

@@

\begin{frame}{Example of a Continuous Quantum Walk}

\begin{frame}{Example of a Discrete-Time Quantum Walk}

\begin{frame}{Continuous vs Discrete Quantum Walks }

\begin{frame}{An Example of a Cayley Graph}

        The adjacency matrix of the Cayley graph $\Gamma = (V,E)$ can be expressed as:
        \[ A = \sum_{a \in G} \lambda_a \ket{\hat{a}} \bra{\hat{a}}, \]
        where $\ket{\hat{a}} = \operatorname{QFT_G} \ket{a}$ is the \textit{Quantum Fourier transform (QFT)} of $\ket{a}$. \\
         \vspace{0.5cm}

        \vspace{2cm}

        ... But, What is QFT??
    \end{textblock*}

\begin{frame}{Quantum Fourier Transform (QFT)}

    Let $G$ be an abelian group. The set of characters of $G$, denoted by $\hat{G}$, is the set of homomorphisms $\chi(a, \cdot): G \to \mathbb{C}$ where $a \in G$. If $G \cong \mathbb{Z}_{N_1} \oplus \cdots \oplus \mathbb{Z}_{N_k}$ then the character $\chi(a, \cdot)$ can be explicitly written as
    \[ \chi(a, x) = \omega_{N_1}^{a_1x_1} \cdots \omega_{N_k}^{a_kx_k} \]

    where $\omega_M = \exp(2\pi i/ M)$ is a primitive $M$-th root of unity. The Fourier transform of a function $f: G \to \mathbb{C}$ is given by
    \[ \hat{f}(a) = \frac{1}{\sqrt{|G|}} \sum_{x \in G} \chi(a, x) f(x). \]
    The quantum Fourier transform:

    \[
        \sum_{g \in G} f(g) \ket{g} \mapsto \sum_{x \in G} \hat{f}(x)\ket{x}
    \]

        The adjacency matrix of the Cayley graph $\Gamma = (V,E)$ can be expressed as:
        \[ A = \sum_{a \in G} \lambda_a \ket{\hat{a}} \bra{\hat{a}}, \]
        Where $\ket{\hat{a}} = \operatorname{QFT}_G(\ket{a}) =  \frac{1}{\sqrt{|G|}}\sum_{g \in G}\chi(a,g)\ket{g}$. \\
        The eigenvalues $\lambda_a$ are given by:
        \[ \lambda_a = \sum_{q \in Q} \chi(a, q). \]
        \begin{itemize}
            \item  Note that the eigenvectors $\ket{\hat{a}}$ of $A$ depend only on $G$ and not on the set $Q$.
        \end{itemize}

    Given a regular group action $(G, X, *)$ with a fixed element $x \in X$ and a set  $Q = \{q_1, q_2, \dots, q_k\} \subset G$, let $\Gamma = (X, E)$ be a graphs with vertex set $X$ and edge set consisting of pairs $(x, y) \in X \times X$ such that $y = q * x$ for some $q \in Q$. The adjacency matrix of $\Gamma$ is
    \[
    A = \sum_{h \in G} \lambda_h \ket{G^{(h)} * x} \bra{G^{(h)} * x}, 
    \]
    Where:
    \[
        \ket{G^{(h)} * x} = \frac{1}{\sqrt{|G|}}\sum_{g \in G}\chi(g,h)\ket{g*x}
    \]
    And $\lambda_h = \sum_{q \in Q} \chi(h, q)$. 
    \vspace{0.5cm}
    \begin{itemize}
        \item Again, the eigenvectors $\ket{G^{(h)} * x}$ depend only on $G$.
    \end{itemize}
    \end{textblock*}
\end{frame}

% \begin{frame}{Simulating group action quantum walks}

%         \[
%         e^{-iAt} = \sum_{h \in G}\ket{G^{(h)}*x}\bra{G^{(h)}*x}
%         \]

%         \[
%         T := \sum_{y \in X}\sum_{b \in {\{0,1\}}}(\ket{j}\bra{j}\otimes\ket{b}\bra{b})\otimes\ket{\phi_{yb}}
%         \]

%         \[
%         = \sum_{y \in X}\sum_{b \in {\{0,1\}}}\ket{y}\ket{b}\ket{\phi_{yb}}\bra{y}\bra{b}
%         \]

%         \[
%             \ket{\phi_{y0}} = \frac{1}{\sqrt{|Q|}}\sum_{q \in Q}\ket{q*y}\ket{0}
%         \]

    \begin{textblock*}{12cm}(0.5cm,1.5cm)
        \textit{Hartley Transform:}\\
        item Let $N$ be a positive integer, and let $\mathbb{Z}_N$ be the additive cyclic group of integers modulo $N$. The Hartley transform of a function $f: \mathbb{Z}_N \to \mathbb{R}$ is the function $\cht_N(f): \mathbb{Z}_N \to \mathbb{R}$ defined by
        \[ \cht_N(f)(a) = \frac{1}{\sqrt{N}} \sum_{y = 0}^{N - 1} \cas\Big(\frac{2 \pi ay}{N}\Big) f(y),  \]
        where $\cas(x) = \cos(x) + \sin(x)$ \\
        For a single basis element of the cyclic group $\mathbb{Z}_N$, the quantum Hartly transform simplifies to
        \begin{equation}
            \label{eq:qht-N}
            \qht_N: \ket{a} \mapsto \frac{1}{\sqrt{N}} \sum_{y = 0}^{N - 1} \cas\Big(\frac{2\pi a y}{N}\Big) \ket{y}.
        \end{equation}
    \end{textblock*}

    \begin{textblock*}{12cm}(0.5cm,1.5cm)
        Lemma: The money state $\ket{\mathbb{Z}_N^{(h)} * x}_H$ is an eigenstate of $W$ with eigenvalue $e^{i\lambda_h t}$.
        % \vspace{1cm}

        \textit{Proof:}
        \begin{align*}
            e^{iAt} \ket{\mathbb{Z}_N^{(h)} * x}_H
            & = \sum_{g \in \mathbb{Z}_N} e^{i\lambda_gt} \ket{\mathbb{Z}_N^{(g)} * x}\braket{\mathbb{Z}_N^{(g)} * x|\mathbb{Z}_N^{(h)} * x}_H 
        \end{align*}

        \[
        =\sum_{g \in \mathbb{Z}_N} e^{i\lambda_gt} \ket{\mathbb{Z}_N^{(g)} * x} \bra{\mathbb{Z}_N^{(g)} * x} \Big( \frac{1 - i}{2} \ket{\mathbb{Z}_N^{(h)} * x} + \frac{1 + i}{2} \ket{\mathbb{Z}_N^{(-h)} * x} \Big) 
        \]

        \[
        =e^{i\lambda_{h}t} \frac{1 - i}{2} \ket{\mathbb{Z}_N^{(h)} * x} + \frac{1 + i}{2} e^{i\lambda_{-h}t} \ket{\mathbb{Z}_N^{(-h)} * x} 
        \]

        \[
        =e^{i\lambda_{h}t} \ket{\mathbb{Z}_N^{(h)} * x}_H,
        \]

        where the last equality follows from the fact that $\lambda_h = \lambda_{-h}$.
    \end{textblock*}

        \vspace{1cm}
        If we choose $t = \operatorname{poly}(\log N)$, it follows from Lemma that we can run the \textit{phase estimation algorithm} with the unitary $W$ and the eigenstate $\ket{\mathbb{Z}_N^{(h)} * x}_H$ to compute an estimate $\tilde{\lambda}_h$ of $\lambda_h$ such that %$\abs{\tilde{\lambda}_h - \lambda_h} \le 1 / poly(log N)$. 
        % \operatorname{poly}(\log N)
        \[
        |\tilde{\lambda}_h - \lambda_h| \leq \frac{1}{\operatorname{poly}(\log N)}
        \]

@@

    \begin{textblock*}{12cm}(0.5cm,1.5cm)
        Now, let us briefly explain how the algorithm for $\qft_N$ works:
        \begin{align*}
            \qft_N\ket{a}
            & = \frac{1}{\sqrt{N}} \sum_{y = 0}^{N - 1} \omega_N^{ay}\ket{y} \\
            & = \frac{1}{\sqrt{N}} \sum_{y = 0}^{N / 2 - 1} \omega_N^{ay} \ket{y} + (-1)^a \sum_{y = 0}^{N / 2 - 1} \omega_N^{ay} \ket{y + N/2} \\
            & = \frac{1}{\sqrt{N / 2}} \sum_{y = 0}^{N / 2 - 1} \omega_N^{ay} \frac{1}{\sqrt{2}} (\ket{0} + (-1)^a \ket{1}) \ket{y}, \numberthis\label{eq:qft_alt}
        \end{align*}

@@

\begin{frame}{References}

@@
